---
title: "Basic Webscraping"
---

<style type="text/css">
.table {

    width: 80%;
    margin-left:10%; 
    margin-right:10%;
}
</style>
```{r,setup, echo=FALSE, cache=TRUE}
## numbers >= 10^5 will be denoted in scientific notation,
## and rounded to 2 digits
options(scipen = 3, digits = 3)
```



## Exercises

1. Read the HTML content of the following URL with a variable called webpage: https://money.cnn.com/data/us_markets/ At this point, it will also be useful to open this web page in your browser.

```{r}
library(rvest)
library(httr)

site_url <- "https://money.cnn.com/data/us_markets/"
webpage <- read_html(site_url)
```

2. Get the session details (status, type, size) of the above mentioned URL.

```{r}
html_session(site_url)

```

3. Extract all of the sector names from the “Stock Sectors” table (bottom left of the web page.)

```{r}

# "Stock Sectors" css identified with SelectorGadget tool. Extract appropriate nodes and their associated text values.

stock_sectors <- html_text(html_nodes(webpage, "#wsod_sectorPerformance .wsod_firstCol"))

stock_sectors
```

4. Extract all of the “3 Month % Change” values from the “Stock Sectors” table.

```{r}

# "3 Month % Change" css identified with SelectorGadget tool. Extract appropriate nodes and their associated text values.

three_change <- html_text(html_nodes(webpage, "#wsod_sectorPerformance td.wsod_aRight"))

three_change
```

5. Extract the table “What’s Moving” (top middle of the web page) into a data-frame.

```{r}

# Select nodes identified as tables and use html_table function which converts the tables to a data frame. Here we are only interested in the first table, the "What's Moving" table, as indicated by the [1] index.

what_moving <- html_table(html_nodes(webpage, "table")[1])

what_moving 
```

6. Re-construct all of the links from the first column of the “What’s Moving” table.
Hint: the base URL is “https://money.cnn.com”

```{r}
# Column css identified with the SelectorGadget tool. Only the first ten rows are part of the "What's Moving" table, the remaining values of the column are actually part of the "Stock Sectors" table
what_moving_nodes <- html_nodes(webpage, ".wsod_firstCol")[1:10]

# Select the nodes labelled as "a" for hyperlinks and extract the "href" attribute containing the link fragment associated with that item
what_moving_link_frag <- html_nodes(what_moving_nodes, "a") %>% 
  html_attr("href")

base_url <- "https://money.cnn.com"

# Attach the link fragments to the base url to create working, complete links
what_moving_full_links <- paste(base_url, what_moving_link_frag, sep='')

what_moving_full_links 

```

7. Extract the titles under the “Latest News” section (bottom middle of the web page.)

```{r}
# Latest news css identified with SelectorGadget tool

latest_news <- html_text(html_nodes(webpage, "#section_latestnews li"))

latest_news
```

8. To understand the structure of the data in a web page, it is often useful to know what the underlying attributes are of the text you see.
Extract the attributes (and their values) of the HTML element that holds the timestamp underneath the “What’s Moving” table.

```{r}

# Timestamp css identified with the SelectorGadet tool

timestamp <- html_node(webpage, ".wsod_disclaimer span") %>% html_attrs()

timestamp

```

9. Extract the values of the blue percentage-bars from the “Trending Tickers” table (bottom right of the web page.)
Hint: in this case, the values are stored under the “class” attribute.

```{r}
# Extract the nodes that have class "bars" and select the class attribute
blue_bars <- html_nodes(webpage, ".bars") %>% html_attr("class")

# Each of the values is preceded by "bars pct" followed by the percentage. Remove these characters and convert to a number. 
bar_nums <- as.numeric(sub("bars pct", '', blue_bars))

bar_nums
```

10. Get the links of all of the “svg” images on the web page.

```{r}
# Used the Ctrl+F function on the Inspect page of the webpage source to find all instances of "svg", the images associated with "svg" were all of class="logo-cnn"

svg_link_frag <- html_nodes(webpage, ".logo-cnn") %>% html_attr("src")
svg_full_link <- paste(base_url, svg_link_frag, sep='')

svg_full_link
```

