# pre-class

# test commit

Make sure you commit this often with meaningful messages. 


### Getting Started


We will work with the dataset called [gapminder](https://github.com/jennybc/gapminder), this is a cleaned up version from [Gapminder Data](http://www.gapminder.org/data/). Gapminder contains a lot of great data on all of the nations of the world. We first need to install the gapminder package in R. 

```{r}
install.packages("gapminder")
```

```{r}
library(dplyr)
library(gapminder)
gapminder

glimpse(gapminder)
```



### Pre-Class Problems

Use **dplyr** functions to address the following questions:

1. How many unique countries are represented per continent?
```{r}
group_by(gapminder, continent) %>%
summarise(noOfCountries <- n_distinct(country))
```

2. Which European nation had the lowest GDP per capita in 1997? 

```{r}
lowGdp97 <- filter(gapminder, continent == "Europe", year == "1997") %>%
  arrange(gdpPercap)

print(lowGdp97[1,1])
```

3. According to the data available, what was the average life expectancy across each continent in the 1980s?

```{r}
group_by(gapminder, continent) %>%
  filter(year %in% c(1980:1989)) %>%
  summarise(contLifeExp <- mean(lifeExp))

```

4. What 5 countries have the highest total GDP over all years combined?
```{r}
glimpse(gapminder)

highGDP <- group_by(gapminder, country) %>%
  summarise(totalGDP = sum(gdpPercap)) %>%
  arrange(desc(totalGDP))

highGDP[1:5, 1]
```

5. What countries and years had life expectancies of _at least_ 80 years? _N.b. only output the columns of interest: country, life expectancy and year (in that order)._

```{r}
select(gapminder, country, lifeExp, year) %>%
  filter(lifeExp >= 80)
  
```

6. What 10 countries have the strongest correlation (in either direction) between life expectancy and per capita GDP?
```{r}
group_by(gapminder, country) %>%
  mutate(lifeGDPcor = abs(cor(lifeExp, gdpPercap))) %>%
  summarise(correlation = mean(lifeGDPcor)) %>%
  arrange(desc(correlation))

```

7. Which combinations of continent (besides Asia) and year have the highest average population across all countries? _N.b. your output should include all results sorted by highest average population_. With what you already know, this one may stump you. See [this Q&A](http://stackoverflow.com/q/27207963/654296) for how to `ungroup` before `arrange`ing. This also [behaves differently in more recent versions of dplyr](https://github.com/hadley/dplyr/releases/tag/v0.5.0).

```{r}
group_by(gapminder, continent, year) %>%
  filter(continent != "Asia") %>%
  summarize(avgPop = mean(pop)) %>%
  ungroup() %>%
  arrange(desc(avgPop))
```

8. Which three countries have had the most consistent population estimates (i.e. lowest standard deviation) across the years of available data? 

```{r}
group_by(gapminder, country) %>%
  summarise(sd = sd(pop)) %>%
  arrange(sd)

# Sao Tome and Principe, Iceland, Montenegro
```

9. Subset **gm** to only include observations from 1992 and store the results as **gm1992**. What kind of object is this?
10. Which observations indicate that the population of a country has *decreased* from the previous year **and** the life expectancy has *increased* from the previous year? See [the vignette on window functions](https://cran.r-project.org/web/packages/dplyr/vignettes/window-functions.html).
